Preliminary Insights into Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan

### 2025-06-07 analysis/paper/paper.qmd
revise based on comments

Deer bone surface modifications relative to human butchering and cooking practices, such as cut marks, fracturing, and burning damage, were recorded, along with natural marks like weathering and root etching to understand taphonomic effects. Cut marks on major limb joints, including the shoulder, elbow, wrist, hip, knee, and ankle [@Lyman2005], were quantified to explore butchering patterns over time. Moreover, cut marks on lower limbs related to skinning processes, were examined and compared. Such skinning marks are characteristic of circular and straight incisions left by a series of cuts for acquiring intact animal hides [@Costamagno2019; @Isaakidou2007].

>>>>

Deer bone surface modifications relative to human butchering and cooking practices, such as cut marks, fracturing, and burning damage, were recorded, along with natural marks like weathering and root etching to understand taphonomic effects. Cut marks on major limb joints, including the shoulder, elbow, wrist, hip, knee, and ankle [@Lyman2005], were quantified to explore butchering patterns over time. Moreover, cut marks on lower limbs were particularly examined since they might be related to skinning processes. During skinning, circular and longitudinal incisions are required to obtain intact animal hides that often leave transverse, oblique, or longitudinal marks on lower limb [@Costamagno2019; @Isaakidou2007; @Val2011].
